---
title: "Practical 8"
author: "Functional Programming"
date: "Jodi Tay Seow Xuan"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Write your own my_cumsum() function

The base R function cumsum() returns the cumulative sum of its argument, e.g.

```{r}
cumsum(1:4)
```

  a. Write the function my_cumsum() that does the same.
  
```{r}
my_cumsum <- function(x) {
  
  output <- numeric(length(x))
  
  for (i in 1:length(x)) {
    
     output[i] <- sum(x[1:i])
  }
  
  return(output)
}
```
  
  b. Show that the function works for the argument -1:4.

```{r}
my_cumsum(-1:4)
```

# Bootstrap

a. Source the bootstrap() function of the lecture slides.

```{r}
library(mosaic)
```

  b. Fit the linear model mpg ~ disp + drat + qsec of the mtcars data and save the lm object as fit.

```{r}
data(mtcars)
fit <- lm(mpg ~ disp + drat + qsec, data = mtcars)
```

  c. Display the summary of this model.

```{r}
(sum.fit <- summary(fit))
```

  d. Save the bootstrap object for the same model.

```{r}
library(mosaic)
boot_fit <- do(1000)*lm(mpg ~ disp + drat + qsec, data = mosaic::resample(mtcars))
```

  e. Make an object with the bootstrap confidence intervals and one with the theoretical confidence intervals, and compare both.

```{r}
confint(boot_fit, level = 0.95)
confint(fit)
```

```{asis}
The bootstrap confidence interval and theoretical confidence interval are quite similar for the variables `disp` and `qsec` and the intercept, suggesting that the estimates obtained from the resampling and theoretical sampling are consistent with each other.

However, the bootstrap confidence interval is significantly wider than the theoretical confidence interval for `drat`, suggesting a higher level of uncertainty in the resampling estimate.
```

  f. In a 2 x 2 plot array, plot the bootstrap sampling distributions of the four parameters in red, and the theoretical sampling distribution of the parameters in blue. For the latter, use stat_function(fun = dnorm, args = list(mean = ..., sd = ...)), using the mean and sd values of the corresponding lm model estimates.

```{r}
library(ggplot2)
library(gridExtra)
mpg_plot <- ggplot(boot_fit) + 
  geom_density(aes(x = Intercept), col = "red") +
  stat_function(fun = dnorm,
                args = list(
                  mean = 11.524390,
                  sd = 11.887430),
                col = "blue")

disp_plot <- ggplot(boot_fit) + 
  geom_density(aes(x = disp), col = "red") +
  stat_function(fun = dnorm,
                args = list(
                  mean = fit$coef[2],
                  sd = sum.fit$coef[2, 2]),
                col = "blue")

drat_plot <- ggplot(boot_fit) + 
  geom_density(aes(x = drat), col = "red") +
  stat_function(fun = dnorm,
                args = list(
                  mean = fit$coef[3],
                  sd = sum.fit$coef[3, 2]),
                col = "blue")

qsec_plot <- ggplot(boot_fit) + 
  geom_density(aes(x = qsec), col = "red") +
  stat_function(fun = dnorm,
                args = list(
                  mean = fit$coef[4],
                  sd = sum.fit$coef[4, 2]),
                col = "blue")

grid.arrange(mpg_plot, disp_plot, drat_plot, qsec_plot, ncol = 2)
```

---

End of practical