---
title: "Practical 3"
author: "Jodi Tay Seow Xuan"
date: "2024-02-28"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
    number_sections: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary statistics

Make a data frame df with 100 observations, row names id.1 to id.100, and 4 variables (without making the the separate vectors first):

+ a random Poisson variable Y with mean 100

+ a random normal variable X with mean 100 and standard deviation 10

+ a factor S containing the value “f” repeated 40 times and the value “m” repeated 60 times

+ a character vector R containing the letters “a” to “d”, each occuring 25 times

```{r}
df <- data.frame(Y = rpois(100, 100),
                 X = rnorm(100, 100, 10),
                 S = factor(rep(c("f", "m"), c(40, 60))),
                 R = rep(letters[1:4], each = 25),
                 row.names = paste("id", 1:100, sep = "."))
```

Display the mean of X.

```{r}
mean(df$X)
```

Display the standard deviation of Y.

```{r}
sd(df$Y)
```

Display the frequencies of the levels of S.

```{r}
table(df$S)
```

Display the correlation of X and Y.

```{r}
cor(df$X, df$Y)
```

Display the covariance matrix of X and Y (use only the x argument in the cov() function).

```{r}
cov(df[, 1:2])
```

## Logical tests

Display a logical vector with value TRUE if X is greater than 110 and FALSE otherwise.

```{r}
df$X > 110
```

Display the vector from the previous exercise as a numeric vector.

```{r}
as.numeric(df$X > 110)
```

Display the row numbers of df for which Y has a value smaller than 90.

```{r}
which(df$Y < 90)
```

Display the row names of df for which Y has a value smaller than 90.

```{r}
rownames(df)[which(df$Y < 90)]
```

Test whether the variable R is a factor.

```{r}
is.factor(df$R)
```

Convert the variable R into a factor (overwrite the original version in df), and test if the conversion has been successful.

```{r}
df$R <- as.factor(df$R)
is.factor(df$R)
```

Display the first 5 values of the variable X

```{r}
head(df$X, 5)
```

Convert the class of the variable X from numeric to integer (again, overwrite the version in df), and display the first 5 values of X.

```{r}
df$X <- as.integer(df$X)
head(df$X, 5)
```

## Functions with formulas

Display the means of Y within the levels of S.

```{r}
aggregate(Y ~ S, df, mean)
```

Display the variances of Y within the levels of R.

```{r}
aggregate(Y ~ R, df, var)
```

Display the frequency table for the variables S and R

```{r}
xtabs(~ S + R, df)
```

Display the outcome of a t-test with Y as dependent variable and S as grouping factor.

```{r}
t.test(Y ~ S, df)
```

Display the summary of an ANOVA with X as dependent variable and R as grouping factor.

```{r}
summary(aov(X ~ R, df))
```

Display Tukey’s HSD pairwise comparisons for the above ANOVA.

```{r}
TukeyHSD(aov(X ~ R, df))
```

Do the same pairwise comparison with a function starting with the letters pair.

```{r}
pairwise.t.test(df$X, df$R)
```

## Using help

For the following exercises you need to find the appropriate function (these functions have not been discussed in the lecture sheets). For each of the exercises, use a single line of code and as few functions/operators as possible!

Display the minimum value of the variable Y.

```{r}
min(df$Y)
```

Display the row number of the observation with the largest value X value.

```{r}
which.max(df$X)
```

Display a vector with values of Y ordered from small to large.

```{r}
sort(df$Y)
```

Display the classes of the variables in df with a single function call.

```{r}
sapply(df, class)
```

Test the correlation between X and Y for statistical significance.

```{r}
cor.test(df$X, df$Y)
```

Convert the factor R from a nominal variable to an ordinal variable (overwrite the version in df), and display the content of the variable.

```{r}
df$R <- as.ordered(df$R) 
df$R
```

Display the first 6 observations of R.

```{r}
head(df$R, 6)
```

When finished, submit your Rmd file to Blackboard before the deadline.

---

END OF PRACTICAL